A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-02 tex/codatadp.tex
fix

* $\sigma, \tau$ range over types---whether (co)datatypes or uninterpreted types
* $\delta$ range over (co)datatypes

@@

* function symbols are accompanied by a type signature, e.g. $\const{f} : \sigma_1 \times \cdots \times \sigma_n \to
\sigma$ indicates

* don't distinguish between constants and function symbols
  a constant is merely the case $n = 0$ of a function $f :

@@

A (co)datatype specification consists of $l$~mutually recursive types which are
either all datatypes or all codatatypes. Each datatype $\delta$ is equipped by
$m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
arguments and returns a $\delta$ value. The argument types must be either
plain, be or be one of the newly introduced (co)datatypes. To every
argument corresponds a selector. The introduced names for the (co)datatypes, the
constructors, and the selectors must be distinct and different from already
introduced ones. Schematically:
%
\[
\begin{aligned}[t]
(\keyw{co})\keyw{datatype}\;\,
  \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \sigma_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \sigma_{11n_{11}}) & \\
           & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \sigma_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \sigma_{1mn_{1m}}) \\[-1\jot]
   \smash{\vdots\,\,\,} \\[-1\jot]
  \keyw{and}\; \,\delta_l & =\, \ldots
\end{aligned}
\]
%
with
$\const{C}_{ij} : \sigma_{ij1}\times\cdots\times\sigma_{ijk_{ij}} \to \delta_i$
and $\const{s}_{ijk} : \delta_i \to \sigma_{ijk}$. Defaults are assumed for
the selector names if they are omitted.

>>>>

* $\tau$ range over types---whether (co)datatypes or uninterpreted types
* $\delta$ range over (co)datatypes

@@

* function symbols are accompanied by a type signature, e.g. $\const{f} : \tau_1 \times \cdots \times \tau_n \to
\tau$ indicates

* don't distinguish between constants and function symbols
  a constant is merely the case $n = 0$ of a function $f : $

@@

A (co)datatype specification consists of $l$~mutually recursive types which are
either all datatypes or all codatatypes. Each datatype $\delta$ is equipped by
$m \ge 1$ constructors, and each constructor for $\delta$ takes zero or more
arguments and returns a $\delta$ value. The argument types must be either
plain, be or be one of the newly introduced (co)datatypes. To every
argument corresponds a selector. The introduced names for the (co)datatypes, the
constructors, and the selectors must be distinct and different from already
introduced ones. Schematically:
%
\[
\begin{aligned}[t]
(\keyw{co})\keyw{datatype}\;\,
  \delta_1 & {}=\, \const{C}_{11}(\bigl[\const{s}_{111}{:}\bigr]\; \tau_{111}, \,\ldots,\, \bigl[\const{s}_{11n_{11}}{:}\bigr]\; \tau_{11n_{11}}) & \\
           & {}\;\phantom{=\,}\;{} \smash{\,\,\,\vdots} \\[-\jot] % {}\;\phantom{=}\llap{\ensuremath{\mid}}\; \cdots \\
           & {}\;\phantom{=\,}\llap{\ensuremath{\,\mid\,}}\; \const{C}_{1m}(\bigl[\const{s}_{1m1}{:}\bigr]\; \tau_{1m1}, \,\ldots,\, \big[\const{s}_{1mn_{1m}}{:}\big]\; \tau_{1mn_{1m}}) \\[-1\jot]
   \smash{\vdots\,\,\,} \\[-1\jot]
  \keyw{and}\; \,\delta_l & =\, \ldots
\end{aligned}
\]
%
with
$\const{C}_{ij} : \tau_{ij1}\times\cdots\times\tau_{ijk_{ij}} \to \delta_i$
and $\const{s}_{ijk} : \delta_i \to \tau_{ijk}$. Defaults are assumed for
the selector names if they are omitted.
